---
title: "Morphometric_Quantification_RAgas"
output: html_document
date: "2024-01-02"
---

```{r}
library(tidyverse)
library(readxl)
library(cowplot)
library(tidyr)
library(reshape2)
library(viridis)
library(scico)
library(RColorBrewer)



# Theme setting
my_theme <- theme_bw(base_size = 12, base_family = "") + 
  theme(panel.background = element_blank(),
                  strip.background = element_blank(),
                  axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, color = "black"),
                  panel.grid.major.x = element_blank(),
        text = element_text(color = "black")) 
                  
  
```


# plot for whole properties
```{r}
My_col <- c("#778899", "#233B6C", "#DDA0DD", "#F8B400",  "#00BFFF","#3CB371")
My_order <- c("NT_NT", "NT_Mat", "NT_MatRA", "RA_NT", "RA_Mat", "RA_MatRA")

#Loading data
whole <- read_excel("Quantification_gastruloid_whole2.xlsx")

whole2 <- whole %>% filter(Cell_number == 4000) %>% mutate(condition = paste(cond0_24h,cond48_120h, sep = "_"))

whole2$condition <- factor(whole2$condition, levels = My_order)


# Whole length
p1 <- 
  ggplot(whole2, aes(x=Time, y = Whole_length, colour = condition, fill = condition,group = condition)) +
  geom_smooth(aes(), alpha = 0.2, method = "loess", span = 2) +
  geom_jitter(width = 1) +
   scale_x_continuous(breaks = seq(0, 120, by = 24)) +
  ylim(c(0,2000)) +
  my_theme +
  scale_color_manual(values =  My_col)+
  scale_fill_manual(values =  My_col) 
  
  

# Average width
p2 <- 
 ggplot(whole2, aes(x=Time, y = Average_width, colour = condition, fill = condition,group = condition)) +
  geom_smooth(aes(), alpha = 0.2, method = "loess", span = 2) +
  geom_jitter(width = 1) +
   scale_x_continuous(breaks = seq(0, 120, by = 24)) +
  ylim(c(0,900)) +
  my_theme +
  scale_color_manual(values =  My_col)+
  scale_fill_manual(values =  My_col)


# Width_Length ratio
p3 <- 
 ggplot(whole2, aes(x=Time, y = L_per_W, colour = condition, fill = condition,group = condition)) +
  geom_smooth(aes(), alpha = 0.2, method = "loess", span = 2) +
  geom_jitter(width = 1) +
   scale_x_continuous(breaks = seq(0, 120, by = 24)) +
  ylim(c(0,8)) +
  my_theme +
  scale_color_manual(values =  My_col)+
  scale_fill_manual(values =  My_col)




# Neural.length_per_Whole.length
whole3 <- whole2 %>% filter(NT_Wh_ratio != "NA")
whole3$NT_Wh_ratio <- as.numeric(whole3$NT_Wh_ratio)

p4 <- 
  ggplot(whole3, aes(x=Time, y = NT_Wh_ratio, colour = condition, fill = condition, group = condition)) +
  geom_smooth(aes(), alpha = 0.2, method = "loess", span = 2) +
  geom_jitter(width = 1) +
   scale_x_continuous(breaks = seq(0, 120, by = 24)) +
  ylim(c(0,100)) +
  my_theme +
  scale_color_manual(values =  My_col) +
  scale_fill_manual(values =  My_col)


# Average_NT_width
whole3 <- whole2 %>% filter(Average_NT_width != "NA")
whole3$Average_NT_width <- as.numeric(whole3$Average_NT_width)

p5 <- 
  ggplot(whole3, aes(x=Time, y = Average_NT_width, colour = condition, fill = condition, group = condition)) +
  geom_smooth(aes(), alpha = 0.2, method = "loess", span = 2) +
  geom_jitter(width = 1) +
  scale_x_continuous(breaks = seq(0, 120, by = 24)) +
  ylim(c(0, 400)) +
  my_theme +
  scale_color_manual(values =  c("#00BFFF","#3CB371","#778899", "#233B6C", "#F8B400",  "#DDA0DD")) +
  scale_fill_manual(values =  c("#00BFFF","#3CB371","#778899", "#233B6C", "#F8B400",  "#DDA0DD"))


# NTlength_NTwidth
whole3 <- whole2 %>% filter(NTlength_NTwidth != "NA")
whole3$NTlength_NTwidth <- as.numeric(whole3$NTlength_NTwidth)



p6 <- 
  ggplot(whole3, aes(x=Time, y = NTlength_NTwidth, colour = condition, fill = condition, group = condition)) +
  geom_smooth(aes(), alpha = 0.2, method = "loess", span = 2) +
  geom_jitter(width = 1) +
  scale_x_continuous(breaks = seq(0, 120, by = 24)) +
  ylim(c(0, 12)) +
  my_theme +
  scale_color_manual(values =  c("#00BFFF","#3CB371","#778899", "#233B6C", "#F8B400",  "#DDA0DD")) +
  scale_fill_manual(values =  c("#00BFFF","#3CB371","#778899", "#233B6C", "#F8B400",  "#DDA0DD"))
    
p_merge <-  cowplot::plot_grid(p1,p2,p3, p4,p5,p6,ncol = 3) 

# save plot
save_plot(plot = p_merge, filename = "Whole_property.pdf", base_height = 8, base_width = 14 )






# NT width
p1 <- ggplot(whole2, aes(x=Time, y = NT_Width_post, group = Time)) +
  geom_boxplot()+ geom_point() +
  ylim(c(0,400)) +
  my_theme +
  scale_color_manual(values =  My_col)

p2 <- ggplot(whole2, aes(x=Time, y = NT_Width_center, group = Time)) +
  geom_boxplot()+ geom_point() +
  ylim(c(0,400)) +
  my_theme +
  scale_color_manual(values =  My_col)

p3 <- ggplot(whole2, aes(x=Time, y = NT_Width_anterior, group = Time)) + 
  ylim(c(0,400)) +
  geom_boxplot()+ geom_point() +
  my_theme +
  scale_color_manual(values =  My_col)

p_NT_width <-  cowplot::plot_grid(p1,p2,p3, ncol = 3) 
p_NT_width
# save plot
save_plot(plot = p_NT_width, filename = "NT_width.pdf", base_height = 4, base_width = 7 )



```

# somite count
```{r}

My_col <- c("#778899", "#233B6C", "#DDA0DD", "#F8B400",  "#00BFFF","#3CB371")
My_order <- c("NT_NT", "NT_Mat", "NT_MatRA", "RA_NT", "RA_Mat", "RA_MatRA")

# Loading data
whole <- read_excel("Quantification_gastruloid_whole2.xlsx")

whole_s <- whole %>% filter(Cell_number == 4000) %>% filter(Time == 72 |  Time == 96 |  Time == 120) %>% mutate(condition = paste(cond0_24h,cond48_120h, sep = "_"))

whole_s$Time <- as.factor(whole_s$Time)

whole_s$condition <- factor(whole_s$condition, levels = My_order)


# Somite count
# Plot boxplot and jitter
p_somite <- ggplot(whole_s, aes(x = Time, y = Somite_pair_count, fill = condition)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(color = condition), position = position_jitterdodge(jitter.width = 0.3, dodge.width = 0.8)) +
  scale_y_continuous(limits = c(0,10), breaks = seq(0, 10, by = 1)) +
  my_theme + 
  theme(panel.grid.minor.y = element_blank(),
        axis.text.x = element_text(angle = 0, hjust = 0.5, vjust = 0.5, color = "black")) + 
  scale_fill_manual(values = My_col) +scale_color_manual(values = My_col)

p_somite

# save plot
save_plot(plot = p_somite, filename = "Whole_somite_count.pdf", base_height = 4, base_width = 6 )


```



# somite bi-lateral
```{r}

My_col <- c("#2E8B57", "#778899" )
My_order <- c("Paired", "Not_paired")

# Loading data
Lat_count <- read_excel("Quantification_Lateral.xlsx")

Lat_count$somite_position <- factor(Lat_count$somite_position, levels = My_order)


# Somite count
# Plot 100% bar plot
Lat_count %>% ggplot(aes(x = Treatment, y = Count, fill = somite_position)) +
  geom_bar(stat = "identity", position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  labs(x = "", y = "Frequency") +
  scale_fill_manual(values = My_col) +
  my_theme + 
  theme(panel.grid.minor.y = element_blank(),
        axis.text.x = element_text(color = "black")
        )

ggsave("Somite_Lateral.pdf", height = 4, width = 3 )




```


# plot for each somite properties (WT)
```{r}
#Loading data
somite <- read_excel("Quantification_somite.xlsx")


# somite length 
  
  p1 <- somite %>%
  dplyr::filter(Time == 120) %>%
  ggplot(aes(x=Somite_position, y = Somite_length, group = Somite_position)) +
  geom_boxplot(aes(fill = Somite_position))+ geom_point() +
  ylim(c(0,200)) +
  my_theme
  
  
  # save plot
save_plot(plot = p1, filename = "somite_length.pdf", base_height = 3, base_width = 6 )



# somite width
p2 <- somite %>%
  dplyr::filter(Time == 120) %>%
  ggplot(aes(x=Somite_position, y = Somite_width, group = Somite_position)) +
  geom_boxplot(aes(fill = Somite_position))+ geom_point() + #facet_wrap(~Time)+
  ylim(c(0,200)) +
  my_theme

save_plot(plot = p2, filename = "somite_width.pdf", base_height = 3, base_width = 6 )



# somite area 
p3 <- somite %>%
  dplyr::filter(Time == 120) %>%
  ggplot(aes(x=Somite_position, y = Somite_area, group = Somite_position)) +
  geom_boxplot(aes(fill = Somite_position))+ geom_point() + # facet_wrap(~Time)+
  #ylim(c(0,200)) +
  my_theme
save_plot(plot = p3, filename = "somite_area.pdf", base_height = 3, base_width = 6 )



```



# plot for each somite properties (LDN)
```{r}
#Loading data
somite <- read_excel("Quantification_LDN_gastruloid.xlsx", sheet = 2)


# somite length 
  
  p1 <- somite %>%
  dplyr::filter(Time == 120) %>%
  ggplot(aes(x=Somite_position, y = Somite_length, group = Somite_position)) +
  geom_boxplot(aes(fill = Somite_position))+ geom_point() +
  ylim(c(0,200)) +
  my_theme + 
  
  
  p1
  
  # save plot
save_plot(plot = p1, filename = "LDN_somite_length.pdf", base_height = 3, base_width = 6 )



# somite width
p2 <- somite %>%
  dplyr::filter(Time == 120) %>%
  ggplot(aes(x=Somite_position, y = Somite_width, group = Somite_position)) +
  geom_boxplot(aes(fill = Somite_position))+ geom_point() + #facet_wrap(~Time)+
  ylim(c(0,200)) +
  my_theme

save_plot(plot = p2, filename = "LDN_somite_width.pdf", base_height = 3, base_width = 6 )



# somite area 
p3 <- somite %>%
  dplyr::filter(Time == 120) %>%
  ggplot(aes(x=Somite_position, y = Somite_area, group = Somite_position)) +
  geom_boxplot(aes(fill = Somite_position))+ geom_point() + # facet_wrap(~Time)+
  ylim(c(0,20000)) +
  my_theme
save_plot(plot = p3, filename = "LDN_somite_area.pdf", base_height = 3, base_width = 6 )

 

```










# TBX6 positive cells
```{r}
#Loading data
T6 <- read_excel("Quantification_TBX6_24h_gastruloid.xlsx")

# Setting x-axis order

T6$Genotype <- factor(T6$Genotype, levels =  c("NTC", "KO_TBX6"))
My_col <- c("#778899",  "#F8B400")


  p1 <- 
    T6 %>%
  ggplot(aes(x=Genotype, y = ratio, group = Genotype)) +
  geom_boxplot(aes(fill = Genotype))+ geom_point() +
  ylim(c(0,100)) +
    scale_fill_manual(values = My_col) +
  ylab("TBX6 (%)") + 
  xlab("") +
  my_theme 


  # save plot
save_plot(plot = p1, filename = "TBX6_ratio_24h.pdf", base_height = 4, base_width = 4 )


```


# somite live imaging (Time course)
```{r}
sm <- read_excel("Somite_number_live_imaging.xlsx")

msm <- melt(sm, key = Replicates, id.vars = "Replicates", variable.name = "Time", value.name = "Somite_number")

msm$Time <- as.numeric(as.character(msm$Time))


ggplot(msm, aes(x= Time, y = Somite_number, group = Replicates, colour = Replicates)) +
  geom_line(position=position_dodge(width=0.5), size = 1, alpha = 0.5) +
  theme_classic(base_size = 12, base_family = "") +
  scale_color_discrete(type = "PuBu") +
  theme(panel.background = element_blank(),
                  strip.background = element_blank(),
                  axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
                  panel.grid.major.x = element_blank())
ggsave("somite_liveimaging.pdf", width = 6, height = 4)



ggplot(msm, aes(x= Time, y = Somite_number, group = Replicates, colour = Replicates)) +
  geom_line(position=position_dodge(width=0.5), size = 1, alpha = 0.5) +
  theme_classic(base_size = 12, base_family = "") +
  theme(panel.background = element_blank(),
                  strip.background = element_blank(),
                  axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
                  panel.grid.major.x = element_blank())

scale_color_manual(values = color_table$Color, drop = F)
```


# somite live imaging (Interval)
```{r}
int <- read_excel("Somite_number_live_imaging.xlsx", sheet = "Interval")

mint <- melt(int, key = Replicates, id.vars = "Replicates", variable.name = "Segments", value.name = "Interval")

ggplot(mint, aes(x= Segments, y = Interval)) +geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(group = Replicates, colour = Replicates), alpha = 0.5) + 
 # scale_color_viridis(discrete = T, option = "magma") +
  
  ylim(c(0,12))+
  theme_classic(base_size = 12, base_family = "") +
  theme(panel.background = element_blank(),
                  strip.background = element_blank(),
                  axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
                  panel.grid.major.x = element_blank())
ggsave("somite_interval.pdf", width = 6, height = 4)


```


# LDN whole property
```{r}
ldn <- read_excel("Quantification_LDN_gastruloid.xlsx")

# Reorder 
ldn$Treatment <- factor(ldn$Treatment, levels = c("Control", "LDN"))
ldn$Somite_pair_count <- as.integer(ldn$Somite_pair_count)

# Color scheme
my_col <- viridis(3, option = "B")[1:2]


# Whole length
p1 <- ggplot(ldn, aes(x=Treatment, y = Whole_length, group = Treatment, fill = Treatment)) +
  geom_boxplot()+ geom_jitter(width = 0.1, height = 0) +
  ylim(c(0,3000)) +
  my_theme + theme(legend.position = "none", axis.title.x = element_blank()) +
  scale_fill_manual(values = my_col) 

p2 <- ggplot(ldn, aes(x=Treatment, y = NT_length, group = Treatment, fill = Treatment)) +
  geom_boxplot()+ geom_jitter(width = 0.1, height = 0) +
  ylim(c(0,2000)) +
  my_theme + theme(legend.position = "none", axis.title.x = element_blank()) +
  scale_fill_manual(values = my_col) 

p3 <- ggplot(ldn, aes(x=Treatment, y = NT_Wh_ratio, group = Treatment, fill = Treatment)) + 
  ylim(c(0,90)) +
  geom_boxplot()+ geom_jitter(width = 0.1, height = 0) +
  my_theme + theme(legend.position = "none", axis.title.x = element_blank()) +
  scale_fill_manual(values = my_col) 


p4 <- ggplot(ldn, aes(x=Treatment, y = Somite_pair_count, group = Treatment, fill = Treatment)) +
  geom_boxplot()+  geom_jitter(width = 0.1, height = 0) +
  scale_y_continuous(limits = c(0, 12), breaks = seq(0, 12, by = 2)) +
  my_theme + theme(legend.position = "none", axis.title.x = element_blank(), ) +
  scale_fill_manual(values = my_col) 

p_merge <-  cowplot::plot_grid(p1,p2,p3,p4, ncol = 4) 

# save plot
save_plot(plot = p_merge, filename = "LDN_morphology.pdf", base_height = 4, base_width = 9 )




# Tukey multiple comparison test
aov(NT_length ~ Treatment, data = ldn) %>% TukeyHSD() 
aov(Somite_pair_count ~ Treatment, data = ldn) %>% TukeyHSD()
aov(Whole_length ~ Treatment, data = ldn) %>% TukeyHSD()
aov(NT_Wh_ratio ~ Treatment, data = ldn) %>% TukeyHSD()
viridis::scale_fill_viridis(discrete = T, option = "B") 

```
# CHIR whole property
```{r}
CHIR <- read_excel("Quantification_CHIR_gastruloid.xlsx")

# Reorder 
CHIR$Treatment <- factor(CHIR$Treatment, levels = c("Control", "CHIR"))
CHIR$Somite_pair_count <- as.integer(CHIR$Somite_pair_count)

# Color scheme
my_col <- viridis(3, option = "B")[1:2]


# Whole length
p1 <- ggplot(CHIR, aes(x=Treatment, y = Whole_length, group = Treatment, fill = Treatment)) +
  geom_boxplot(outlier.shape = NA)+ geom_jitter(width = 0.1, height = 0) +
  ylim(c(0,5000)) +
  my_theme + theme(legend.position = "none", axis.title.x = element_blank()) +
  viridis::scale_fill_viridis(discrete = T, option = "B") 

p2 <- ggplot(CHIR, aes(x=Treatment, y = NT_length, group = Treatment, fill = Treatment)) +
  geom_boxplot(outlier.shape = NA)+ geom_jitter(width = 0.1, height = 0) +
  ylim(c(0,2000)) +
  my_theme + theme(legend.position = "none", axis.title.x = element_blank()) +
  viridis::scale_fill_viridis(discrete = T, option = "B") 

p3 <- ggplot(CHIR, aes(x=Treatment, y = NT_Wh_ratio, group = Treatment, fill = Treatment)) + 
  ylim(c(0,90)) +
  geom_boxplot(outlier.shape = NA)+ geom_jitter(width = 0.1, height = 0) +
  my_theme + theme(legend.position = "none", axis.title.x = element_blank()) +
  viridis::scale_fill_viridis(discrete = T, option = "B") 


p_merge <-  cowplot::plot_grid(p1,p2,p3, ncol = 3) 

# save plot
save_plot(plot = p_merge, filename = "CHIR_morphology.pdf", base_height = 4, base_width = 7 )



```
# KO whole properties
```{r}
#Loading data
whole <- read_excel("Quantification_KO_gastruloid.xlsx")

# Reorder 
whole$Genotype <- factor(whole$Genotype, levels = c("WT", "KO_PAX3", "KO_TBX6"))


whole$Somite_pair_count <- as.integer(whole$Somite_pair_count)

# Whole length
p1 <- ggplot(whole, aes(x = Genotype, y = Whole_length, group = Genotype, fill = Genotype)) +
  geom_boxplot() + geom_jitter(height = 0, width = 0.1) +
  ylim(c(0,2500)) +
  my_theme + theme(legend.position = "none", axis.title.x = element_blank()) +
  viridis::scale_fill_viridis(discrete = T, option = "B") 

p2 <- ggplot(whole, aes(x = Genotype, y = NT_length, group = Genotype, fill = Genotype)) +
  geom_boxplot() + geom_jitter(height = 0, width = 0.1) +
  ylim(c(0,2000))  +
  my_theme +
  theme(legend.position = "none", axis.title.x = element_blank()) +
  viridis::scale_fill_viridis(discrete = T, option = "B") 

p3 <- ggplot(whole, aes(x= Genotype, y = NT_Wh_ratio, group = Genotype, fill = Genotype)) + 
  ylim(c(0,100)) +
  geom_boxplot() + geom_jitter(height = 0, width = 0.1) +
  my_theme +
  theme(legend.position = "none", axis.title.x = element_blank()) +
  viridis::scale_fill_viridis(discrete = T, option = "B")

# Somite count
p4 <- ggplot(whole, aes(x = Genotype, y = Somite_pair_count, group = Genotype, fill = Genotype)) +
  geom_boxplot() + 
  geom_jitter(height = 0, width = 0.1) +
  ylim(c(0,12)) +
  my_theme + 
  theme(legend.position = "none", axis.title.x = element_blank()) + 
  viridis::scale_fill_viridis(discrete = T, option = "B") 

p_merge <-  cowplot::plot_grid(p1,p2,p3,p4, ncol = 4) 


# save plot
save_plot(plot = p_merge, filename = "KO_morphology.pdf", base_height = 4, base_width = 9 )




ggsave("KO_Whole_somite_count.pdf", width = 2, height = 4)


# Tukey multiple comparison test
aov(NT_length ~ Genotype, data = whole) %>% TukeyHSD() 
aov(Somite_pair_count ~ Genotype, data = whole) %>% TukeyHSD()
aov(Whole_length ~ Genotype, data = whole) %>% TukeyHSD()
aov(NT_Wh_ratio ~ Genotype, data = whole) %>% TukeyHSD()





# Length_per_Width
ggplot(whole, aes(x = Genotype, y = L_per_W, group = Genotype, fill = Genotype)) +
  geom_boxplot() + 
  geom_jitter(height = 0, width = 0.1) +
  ylim(c(0,7)) +
  my_theme + theme(axis.title = element_blank()) + 
  ggtitle("Length_per_Width")


# width transition
long_df <-
  whole %>% 
  select(Genotype, Width_10_percent, Width_50_percent, Width_90_percent) %>% 
  pivot_longer(cols = !Genotype, names_to = "position", values_to = "width")

long_df %>% ggplot(aes(x = position, y = width, colour = Genotype)) + 
  geom_boxplot()


```



# Time window LDN neural area
```{r}
My_col <- c("#778899", "#FFD700", "#F8B400",  "#FF8C00")
My_order <- c("NT", "LDN_0_24h", "LDN_24_48h", "LDN_48_120h")



#Loading data
LDN <- read_excel("Quantification_Neural_area_LDN_gastruloid.xlsx")

# Reorder 
LDN$Treatment <- factor(LDN$Treatment, levels = My_order)

LDN <- LDN %>% filter(Time == 120)


ggplot(LDN, aes(x= Treatment, y = Wh_NT_ratio, colour = Treatment, fill = Treatment, group = Treatment)) +
  geom_boxplot(color = "black", outlier.shape = NA) +
  geom_jitter(width = 0.1) +
  ylim(c(0,80)) +
  my_theme +
  scale_color_manual(values =  My_col) +
  scale_fill_manual(values =  My_col)

ggsave("LDN_timewindow_Neural_area.pdf", width = 6, height = 4)

```

# Retinol/Retinal/RA neural area
```{r}
My_col <- c("#778899", "#FFD700", "#F8B400", "#AFEEEE", "#40E0D0","#DDA0DD", "#EE82EE")
My_order <- c("NT", "RA_500nM", "RA_3µM", "Retinol_500nM", "Retinol_3µM", "Retinal_500nM", "Retinal_3µM")

#Loading data
RA <- read_excel("Quantification_Neural_area_retinol_retinal.xlsx")

# Reorder 
RA$Treatment <- factor(RA$Treatment, levels = My_order)

levels(RA$Treatment)

#RA <- RA %>% filter(Time == 120)


ggplot(RA, aes(x= Treatment, y = Wh_NT_ratio, colour = Treatment, fill = Treatment, group = Treatment)) +
  geom_boxplot(color = "black", outlier.shape = NA) +
  geom_jitter(width = 0.1) +
  ylim(c(0,100)) +
  my_theme +
  scale_color_manual(values =  My_col) +
  scale_fill_manual(values =  My_col)

ggsave("RA_retinol_retinal_timewindow_Neural_area.pdf", width = 6, height = 4)
```


# SOX10 angle
```{r}
#Loading data
SOX10 <- read_excel("SOX10_angle.xlsx")

ggplot(SOX10, aes(x = "test", y = Angle)) +
  geom_boxplot(fill = "#DC143C", outlier.shape = NA) +
  geom_jitter(width = 0.2) +
  theme_classic() +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_text(size =18),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size =12)
        ) + 
  scale_y_continuous(breaks = c(0, 90, 180, 270, 360), limits = c(0,360))
  ggsave("SOX10_angle_boxplot.pdf", width = 2, height = 7)

  
ggplot(SOX10, aes(x = "test", y = Angle)) +
  geom_violin(fill = "#DC143C") +
  geom_jitter() +
  ylim(c(0, 360)) +
  theme_classic()
  ggsave("SOX10_angle_violinplot.pdf", width = 2, height = 7)



```


# Epithelialized neural tube and somite
```{r}

# Load data
data_sheet4 <- read_xlsx("Quantification_gastruloid_whole2.xlsx", sheet = "Sheet4")

# Calculate percentages
data_sheet4 <- data_sheet4 %>%
  mutate(Epithelialized_Percent = (Epithelialized / (Epithelialized + Not_epithelialized)) * 100,
         Not_Epithelialized_Percent = 100 - Epithelialized_Percent)

# Melt data for ggplot
melted_data <- melt(data_sheet4, id.vars = c("name", "structure"), 
                    measure.vars = c("Not_Epithelialized_Percent", "Epithelialized_Percent"))  # Notice the order change here

# Plot
ggplot(melted_data, aes(x = name, y = value, fill = variable)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = c("grey", "blue4"), labels = c("Not_epithelialized", "Epithelialized")) +  # Notice the color order change here
  facet_wrap(~structure, scales = "free_x") +
  labs(y = "Percentage (%)", x = NULL) +
  theme_minimal()+
  theme(strip.text = element_text(face = "bold"), axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, color = "black"))


ggsave("epithelialized_somite_NT.pdf", width = 4, height = 4)

```

